PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-08 docs/_tex/index.tex
first part of Laurent's comments

\begin{abstract}
Coastal areas host seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges, drone images
acquired with multi-spectral sensors offer a promising solution. This
study focuses on using drones acquisitions for mapping intertidal
macrophytes, effectively discriminating between seagrass and green
macroalgae. Ten drone flights were conducted at two different altitudes
(12m and 120m) across diverse European habitats in France and Portugal.
Low altitude flights were used to train a Deep Learning classifier based
on Neural Networks to discrimintate among 5 intertidal vegetation
classes. Drone mapping demonstrated an overall accuracy of 94\% across
all the sites and images, covering a total area of 467 000 m². The model
exhibited an accuracy of 96.4\% in identifying seagrass.
\end{abstract}

@@

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using in situ measurments. The acquisition
of field data in intertidal zones is however notoriously challenging.
Intertidal seagrass meadows are only partially exposed during low tide
and can be situated in difficult-to-reach mudflats, potentially leading
to inaccurate and limited estimations with conventional sampling
techniques \citep{nijland2019}. Satellite observations have been proven
effective in complementing in situ sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance measurments, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing
and in situ measurments, due to their ability to provide
spatially-explicit observations at very high spatial resolutions (pixel
size from cm to mm) and still capturing data at multi-spectral
resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most study cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green or red algae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the southern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red algae (Rhodophyceae), brown algae (Phaeophyceae),
and green algae (Chlorophyceae).

@@

\caption{\label{fig-map}Sites observed by drone. A: Gulf of Morbihan, B:
Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent
terrestrial areas whereas darkgreen represent intertidal areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. A structure-from-motion photogrammetry
software (Agisoft Metashape) was used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking was the same for every flight. First, tying key points
were detected inside of each image and between overlapping images in
order to obtain a sparse point cloud. This cloud was cleaned using
reprojection accuracy metric in order to remove noisy points. A dense
point cloud was then produced using a structure from motion algorithm. A
surface interpolation of this dense point cloud was made to obtain a
digital surface model (DSM), used to reconstruct the multispectral
ortho-image. Across all sites, flights were made at two different
altitudes : 12 m or/and 120 m. Low altitude drone flights produce
ortho-images with a very high spatial resolution (8 mm per pixel),
making it simple to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allow to cover large
areas and produced images with a pixel size of 80 mm (Table 1).

@@

\begin{longtable}[c]{ccccc}
\caption{List of drone flight, summarising the date, the altitude and the purpose
of each flight.}\tabularnewline

@@

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{\multirow[c]{-3}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Morbihan}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² used to assess the presence or absence of 5 key classes of
vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm
over the sediment during low tide), Phaeophyceae (brown algae),
Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae
(red algae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded online on the Global Biodiversity Information Facility (GBIF)
platform \citep{BedeGbif}, a public open portal to store and share
biodiversity data. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software (imageJ). For all quadrat, the hyperspectral reflectance
signatures of each vegetation class was measured using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. The methodological scheme
followed in this study is presented in Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-vegetation}Class of vegetation used to train the
Neural Network model and their standardised spectral signature}

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights were utilized for training and
prediction of the Neural Network model, whereas low-resolution flights
were solely employed for prediction purposes.}

@@

\begin{longtable}[c]{ccc}
\caption{Vegetation Classes of the model and the number of pixels used to train
and validate each class}\tabularnewline

@@

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Rhodophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{5,771}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into 7 different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table 1) were used for training
purposes because their 8 mm spatial resolution allowied to avoid
spectral sub-pixel mixing and to accurately differentiate various
vegetation classes. More than 418,000 pixels at 8 mm resolution from the
3 training flights were used to train the model (Table 2). Twenty one
variables were used by the model as predictors: 10 raw spectral bands of
the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm
to 840 nm), the same 10 spectral bands standardized using a min/max
transformation (Equation~\ref{eq-std} ; \citep{Cao2017}) and the
Normalized difference vegetation index (NDVI, Equation~\ref{eq-ndvi}).
Standardisation of spectral bands is used to eliminate the scaling
differences between spectra and to limit the effect of biomass on the
shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

>>>>

\begin{abstract}
Coastal areas host seagrass meadows, which offer crucial ecosystem
services including erosion control and carbon sequestration. However,
these areas are increasingly impacted by human activities, leading to
seagrass decline and habitat fragmentation. In situ surveys,
traditionally performed to monitor these ecosystems face limitations on
temporal and spatial coverage, particularly in intertidal zones,
prompting the use of satellite data within monitoring programs. Yet,
satellite remote sensing struggles with spatial and spectral resolution,
making it difficult to discriminate seagrass from other macrophytes in
highly heterogenous meadows. To address these challenges and reduce the
issue of mlixed pixel, drone images at a very high spatial resolution
offer a promising solution. This study focuses on using drone
acquisitions with a ten spectral bands sensor mirroring those of
Sentinel-2, for mapping intertidal macrophytes and effectively
discriminating between seagrass and green macroalgae. Nine drone flights
were conducted at two different altitudes (12m and 120m) across diverse
European habitats in France and Portugal. Low altitude flights were used
to train a Deep Learning classifier based on Neural Networks to
discrimintate among five taxonomic classes of intertidal vegetation:
Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae), Phaeophyceae
(Brown algae), Rhodophyceae (Red macroalgae) and Bacillariophyceae
(Diatoms). Drone mapping demonstrated an overall accuracy of 94\% across
all the sites and images, covering a total area of 467 000 m². The model
exhibited an accuracy of 96.4\% in identifying seagrass. This work
showed that although they share the same pigmentary composition,
seagrass and green algae can be discriminated using ten bands
multispectral sensor. The algorithm should now be adapted for Sentinel-2
to see if this discrimination is still operational at a coarser spatial
resolution.
\end{abstract}

@@

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using \emph{in situ} measurements. The
acquisition of field data in intertidal zones is however notoriously
challenging. Intertidal seagrass meadows are only partially exposed
during low tide and can be situated in difficult-to-reach mudflats,
potentially leading to inaccurate and limited estimations with
conventional sampling techniques \citep{nijland2019}. Satellite
observations have been proven effective in complementing in situ
sampling, allowing for the near real-time and consistent retrieval of
seagrass EOVs and EBVs over extensive meadows. \citetext{\citealp[
]{Zoffoli2021}; \citealp[ ]{xu2021}; \citealp[
]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance spectra, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing
and in situ measurements, due to their ability to provide
spatially-explicit observations at very high spatial resolutions (pixel
size from cm to mm) and still capturing data at multi-spectral
resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most study cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green, or red macroalgae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the southern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red, brown and green macroalgae.

@@

\caption{\label{fig-map}Location of drone flights in France and
Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites),
C: Ria de Aveiro Lagoon (Three sites). Green represent intertidal
areas.}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in ten spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. A structure-from-motion photogrammetry
software (Agisoft Metashape) was used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking was the same for every flight. First, tying key points
were detected inside of each image and between overlapping images in
order to obtain a sparse point cloud. This cloud was cleaned using
reprojection accuracy metric in order to remove noisy points. A dense
point cloud was then produced using a structure from motion algorithm. A
surface interpolation of this dense point cloud was made to obtain a
digital surface model (DSM), used to reconstruct the multispectral
ortho-image. Across all sites, flights were made at two different
altitudes : 12 m or/and 120 m. Low altitude drone flights produce
ortho-images with a very high spatial resolution (8 mm per pixel),
making it efficient to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allow to cover large
areas and produced images with a pixel size of 80 mm
(Table~\ref{tbl-flights}).

@@

\begin{longtable}[c]{ccccc}

\caption{\label{tbl-flights}List of drone flight, summarising the date,
the altitude and the purpose of each flight. 12 m and 120 m flights have
a spatial resolution of 8 and 80 mm respectively.}

\tabularnewline

@@

\multicolumn{1}{>{}c}{} & \multicolumn{1}{>{}c}{\multirow[c]{-3}{*}{\parbox{0.75in}{\centering \textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Gulf\ of\ Morbihan}}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 1.5pt}>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Duer}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120m}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Validation}}} \\

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² assessing the presence or absence of five key taxonomic
classes of intertidal vegetation : Bacillariophyceae (Unicellular
benthic diatoms forming large biofilms at the sediment surface during
low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (seagrass),
Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae)
(Figure~\ref{fig-vegetation}). Pictures of each quadrat were uploaded
online on the Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}, a public open portal to store and share biodiversity
data. Each photograph was also processed to estimate the percent cover
of each type of vegetation using an image processing software (imageJ).
For all quadrats, the hyperspectral reflectance signatures of each
vegetation class was measured using an ASD FieldSpec HandHeld 2
spectroradiometer, which acquires reflectance between 325 and 1075 nm,
with 1 nm of spectral resolution. Hyperspectral signatures from quadrats
serve dual purposes: they validate the radiometric calibration of drone
data and contribute to error reduction in photo interpretation. The
workflow developped in this study is presented in
Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-vegetation}The five taxonomic classes of vegetation
used to train the Neural Network model and their raw spectral signatures
at the spectral resolution of the Micasense RedEdge Dual MX.}

@@

\caption{\label{fig-workflow}Schematic representation of the workflow.
Diamonds represent input or output data, and rectangles represent Python
processing algorithms. The overall workflow of this study is divided
into two distinct parts based on the spatial resolution of the drone
flights: high-resolution flights (pixel size: 8 mm) were utilized for
training and prediction of the Neural Network model, whereas
lower-resolution flights (pixel size: 80 mm) were solely employed for
prediction purposes. Validation has been performed on both high and low
resolution flights.}

@@

\begin{longtable}[c]{ccc}

\caption{\label{tbl-validationPX}Vegetation Classes of the model and the
number of pixels used to train and validate each class}

\tabularnewline

@@

\multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Rhodophyceae}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{5,771}}} & \multicolumn{1}{>{}c}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{3,515,421}}} \\

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model with 2 hidden layers. The training
pixels were categorized into seven different classes, representing the
various habitats encountered at the different study sites: Sediment,
Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowied to
avoid spectral sub-pixel mixing and to accurately differentiate
vegetation classes. More than 418,000 pixels at 8 mm resolution from the
3 training flights were used to train the model
(Table~\ref{tbl-validationPX}). Twenty one variables were used by the
model as predictors: the ten raw spectral bands of the Micasense RedEdge
Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same
ten spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the Normalized difference
vegetation index (NDVI, Equation~\ref{eq-ndvi}). Standardisation of
spectral bands is used to eliminate the scaling differences between
spectra and to limit the effect of biomass on the shape of the spectra
\citetext{\citealp[ ]{Douay2022}; \citealp{Davies2023}}.

@@

The workflow of this study revolves around two distinct flight heights
(12 m and 120 m, Figure~\ref{fig-workflow}) where ensuring consistency
between reflectances at both heights is crucial. This comparison was
conducted at sites where low and high altitude flights overlapped. The
low altitude flights were resampled to the same spatial resolution and
grid as the high flights using a median resampling method. Reflectance
values were then extracted, and a scatterplot was generated.
Subsequently, the slope of the linear model and the coefficient of
determination (R²) were computed.
